Fusing Filters with Integer Linear Programming

### 2014-04-12 2014betterfusionforfilters/Main.tex
Rewrite abstract to highlight that our main contribution is how to handle filters

% -----------------------------------------------------------------------------
\begin{document}
\title{Better fusion for filters}
\authorinfo{ 
  Amos Robinson$^\dagger$ 
  \and Ben Lippmeier$^\dagger$
  \and Manuel M. T. Chakravarty$^\dagger$
  \and Gabriele Keller$^\dagger$ 
}{
  \vspace{5pt}
  \shortstack{
    $^\dagger$Computer Science and Engineering \\
    University of New South Wales, Australia \\[2pt]
    \textsf{\{amosr,benl,chak,keller\}@cse.unsw.edu.au}
  }
}

@@

\begin{abstract}
Recent work on array fusion demonstrates how to compile a specific class of data flow programs into single imperative loops.
However, until now there has been no satisfactory way to take the operator graph expressing a complete program and split it into fusible subgraphs.
Finding good clusterings of operators is non-trivial, as we must achieve a balance between minimizing memory traffic and loop overhead, while maximizing cache locality.
We show how to formulate operator clustering as an \emph{integer linear programming} problem, and use an external library to find a good solution. 
\end{abstract}

% BL: I've removed uses of the word "optimal" here because we don't have a proper language cost model, and our solution is ultimately based on heuristics. ILP can find  an "optimal" solution for a heuristic cost metric, but this isn't the thing we're really trying to optimize.

>>>>

% -----------------------------------------------------------------------------
\begin{document}

\title{Fusing Filters with Integer Linear Programming}
% Better title?
% \title{Better fusion for filters}

\authorinfo{ 
  Amos Robinson$^\dagger$ 
  \and Ben Lippmeier$^\dagger$
  \and Manuel M. T. Chakravarty$^\dagger$
  \and Gabriele Keller$^\dagger$ 
}{
  \vspace{5pt}
  \shortstack{
    $^\dagger$Computer Science and Engineering \\
    University of New South Wales, Australia \\[2pt]
    \textsf{\{amosr,benl,chak,keller\}@cse.unsw.edu.au}
  }
}

@@

\begin{abstract}
Recent work on array fusion shows how to extract data-flow graphs from functional programs and compile them into efficient imperative loops. However, the compilation process only handles graphs that can be compiled into \emph{single} loops, and most programs require more. Prior work shows how use Integer Linear Programming (ILP) to \emph{cluster} the operators in a general data-flow graph into subgraphs that can be individually fused, but until now this approach did not handle filter-like operators which produce arrays of a different size than the input. We extend the existing ILP approach with support for filters, using an external ILP solver to find good clusterings.
\end{abstract}

### 2014-04-12 2014betterfusionforfilters/section/01-Introduction.tex
Rewrite abstract to highlight that our main contribution is how to handle filters

Recent work on data flow fusion~\cite{lippmeier2013flow} shows how to compile a specific class of data flow programs into \emph{single}, efficient imperative loops.
This process of ``compilation'' is equivalent to performing array fusion on a combinator based functional array program, as per related work on stream fusion~\cite{coutts2007streamfusion} and delayed arrays~\cite{keller2010repa}.
The key benefits of data flow fusion over prior work are that 1) it can fuse programs that use branching data flows where a produced array is consumed by several consumers, and 2) complete fusion into a single loop is guaranteed for all programs that operate on the same size input data, and contain no fusion-preventing dependencies between combinators.

>>>>

Data flow fusion~\cite{lippmeier2013flow} is a technique to compile a specific class of data flow programs into single, efficient imperative loops. This process of ``compilation'' is equivalent to performing array fusion on a combinator based functional array program, as per related work on stream fusion~\cite{coutts2007streamfusion} and delayed arrays~\cite{keller2010repa}.
The key benefits of data flow fusion over prior work are that 1) it can fuse programs that use branching data flows where a produced array is consumed by several consumers, and 2) complete fusion into a single loop is guaranteed for all programs that operate on the same size input data, and contain no fusion-preventing dependencies between combinators.
